---
title: "Analysis"
author: "Quinn Thomas"
date: "2023-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
relative_scores_directory = "../../scores/parquet/site_id=fcre"
relative_forecasts_directory = "../../forecasts/parquet/site_id=fcre"

```

## Figure 3: observations

```{r}
df3 <- open_dataset(relative_scores_directory) %>% 
  filter(depth == 1.5, 
         variable %in% c("temperature", "oxygen", "fdom", "chla")) %>% 
  select(datetime, observation, variable) %>% 
  distinct() %>% 
  collect()
```

```{r}
# Need to add units to ribbon
df3 %>% 
  ggplot(aes(x = datetime, y = observation)) +
  geom_point() +
  facet_wrap(~variable, scale = "free") +
  theme_bw()

```

## Figure 5: Ecosystem States

Ecosystem states (non-DA vs DA)

Some combination of the following

```{r}
df5 <- open_dataset(relative_scores_directory) %>% 
  filter(horizon %in% c(0, 3, 7, 30), 
         depth == 1.5, 
         variable %in% c("temperature", "oxygen", "fdom", "chla", "NO3NO2", "SRP", "TN", "TP", "NIT_nit", "NIT_amm", "PHS_frp", "NH4"),
         #variable %in% c("temperature", "oxygen", "fdom", "chla"),
         model_id %in% c("glmaed_ms_daily", "glmaed_ms_no_da", "glmaed_ms_daily_manual")) |> 
  collect() |> 
  mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
          model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily manual", model_id),
         model_id = ifelse(model_id == "glmaed_ms_no_da", "no da", model_id),
         model_id = factor(model_id, levels=c("daily","daily manual","no da"))) |> 
  mutate(variable = ifelse(variable == "NIT_nit", "NO3NO2", variable),
         variable = ifelse(variable == "NIT_amm", "NH4", variable),
         variable = ifelse(variable == "PHS_frp", "SRP", variable))

```

### 0 day ahead

```{r}
open_dataset(relative_scores_directory) %>% 
  filter(horizon  <= 0, 
         depth == 1.5, 
         variable %in% c("temperature", "oxygen", "fdom", "chla", "NO3NO2", "SRP", "TN", "TP", "NIT_nit", "NIT_amm", "PHS_frp", "NH4"),
         #variable %in% c("temperature", "oxygen", "fdom", "chla"),
         model_id %in% c("glmaed_ms_daily", "glmaed_ms_no_da", "glmaed_ms_daily_manual")) |> 
  collect() |> 
  mutate(variable = ifelse(variable == "NIT_nit", "NO3NO2", variable),
         variable = ifelse(variable == "NIT_amm", "NH4", variable),
         variable = ifelse(variable == "PHS_frp", "SRP", variable)) %>% 
  mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
          model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily manual", model_id),
         model_id = ifelse(model_id == "glmaed_ms_no_da", "no da", model_id),
         model_id = factor(model_id, levels=c("daily","daily manual","no da"))) |> 
  ggplot(aes(x = datetime)) +
  geom_point(aes(y = observation), size = 0.8, color = "black") +
  geom_ribbon(aes(ymin = quantile02.5, ymax = quantile97.5), alpha = 0.5, fill = "lightblue") +
  geom_line(aes(y = median), color = "blue") +
  facet_grid(variable~model_id, scale = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5,vjust = 0.5))
```

### 3 day ahead

```{r}
df5 %>% 
  filter(horizon == 3) %>% 
  ggplot(aes(x = datetime)) +
  geom_point(aes(y = observation), size = 0.8, color = "black") +
  geom_ribbon(aes(ymin = quantile02.5, ymax = quantile97.5), alpha = 0.5, fill = "lightblue") +
  geom_line(aes(y = median), color = "blue") +
  facet_grid(variable~model_id, scale = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5,vjust = 0.5))
```

### 7 day ahead

```{r}
#Add units to variable ribbon
df5 %>% 
  filter(horizon == 7) %>% 
  ggplot(aes(x = datetime)) +
    geom_point(aes(y = observation), size = 0.8, color = "black") +
  geom_ribbon(aes(ymin = quantile02.5, ymax = quantile97.5), alpha = 0.5, fill = "lightblue") +
  geom_line(aes(y = median), color = "blue") +
  facet_grid(variable~model_id, scale = "free_y") +
  labs(x = "date time", y = "forecast") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5,vjust = 0.5))
```

### 30 day ahead

```{r}
df5 %>% 
  filter(horizon == 30) %>% 
  ggplot(aes(x = datetime)) +
  geom_point(aes(y = observation), size = 0.8) +
  geom_ribbon(aes(ymin = quantile02.5, ymax = quantile97.5), alpha = 0.5, fill = "lightblue") +
  geom_line(aes(y = median), color = "blue") +
  facet_grid(variable~model_id, scale = "free_y") +
  theme_bw() +
  labs(x = "date time", y = "forecast") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5,vjust = 0.5))
```

## Figure 6: Horizon Synthesis

Different horizons all using daily DA (daily, weekly, fortnightly, monthly)

```{r}
df6 <- open_dataset(relative_scores_directory) %>% 
  filter(depth == 1.5, 
         horizon >= 0, 
         variable %in% c("temperature", "oxygen", "fdom", "chla"),
         model_id %in% c("glmaed_ms_daily", "glmaed_ms_no_da", "glmaed_ms_daily_manual"),
         model_id != "climatology") %>% 
  collect() 
```

```{r}
df6 %>% 
  group_by(variable, model_id) %>% 
  mutate(mean_obs = mean(observation, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(error = abs(median - observation),
         denom = abs(mean_obs - observation)) %>% 
  group_by(variable, horizon, model_id) %>% 
  summarize(rel_abs_error = sum(error, na.rm = TRUE)/sum(denom, na.rm = TRUE),
            logs = median(logs, na.rm = TRUE),
            coff_variation = median(sd, na.rm = TRUE) / median(mean, na.rm = TRUE)) %>% 
  pivot_longer(rel_abs_error:coff_variation, names_to = "metric", values_to = "value") %>% 
  mutate(metric = ifelse(metric == "rel_abs_error", "relative\nabsolute error", metric),
         metric = ifelse(metric == "coff_variation", "coefficient\nof variation", metric),
         metric = factor(metric, levels=c("relative\nabsolute error","coefficient\nof variation","logs"))) %>% 
  mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
          model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily manual", model_id),
         model_id = ifelse(model_id == "glmaed_ms_no_da", "no da", model_id),
         model_id = factor(model_id, levels=c("daily","daily manual","no da"))) |> 
  ggplot(aes(x = horizon, y = value, color = variable)) +
  geom_line() +
  facet_grid(metric~model_id, scales = "free_y") +
  theme_bw() +
  labs(x = "Horizon (days in future)", y = "metric value")
```
## Other figures

### Skill score
```{r}

df7 <- open_dataset(relative_scores_directory) %>% 
  filter(depth == 1.5, 
         horizon >= 0, 
         variable %in% c("temperature", "oxygen", "fdom", "chla"),
         model_id %in% c("climatology","glmaed_ms_daily", "glmaed_ms_daily_manual", "glmaed_ms_no_da")) %>% 
  collect() %>% 
     mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
            model_id = ifelse(model_id == "glmaed_ms_no_da", "no_da", model_id),
            model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily_manual", model_id))
```

Skill score version 1: mean crps by horizon -> mean skill score by horizon

```{r}
df7 %>% 
  select(reference_datetime, horizon, variable, crps, model_id) %>% 
  summarise(crps = median(crps, na.rm = TRUE), .by = c(horizon, variable, model_id)) %>% 
  pivot_wider(names_from = model_id, values_from = crps) %>% 
  mutate(`daily` = 1 - (daily/climatology),
         `no da` = 1 - (no_da/climatology),
         `daily manual` = 1 - (daily_manual/climatology)) %>%
  select(horizon, `daily`,`no da`,`daily manual`, variable) |> 
  pivot_longer(`daily`:`no da`, names_to = "simulation", values_to = "skill") %>% 
  ggplot(aes(x = horizon, y = skill, color = simulation)) + 
  geom_line() +
  facet_wrap(~variable)  +
  theme_bw()
```
### Example forecast


```{r}
ref_datetimes <- open_dataset(relative_scores_directory) %>% 
  filter(depth == 1.5, 
         horizon == 1, 
         variable %in% c("temperature")) %>% 
  distinct(reference_datetime) %>% 
  collect() %>% 
  arrange(reference_datetime) %>% 
  pull(reference_datetime)


```

```{r}
open_dataset(relative_forecasts_directory) %>% 
  filter(depth %in% c(1.5),  
         reference_datetime == ref_datetimes[73], 
         variable %in% c("temperature", "oxygen", "fdom", "chla", "PHY_green", "PHY_cyano", "PHY_diatom")) %>% 
  select(datetime, depth, model_id, variable, prediction, parameter, reference_datetime) %>% 
  collect() %>% 
  mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
          model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily manual", model_id),
         model_id = ifelse(model_id == "glmaed_ms_no_da", "no da", model_id),
         model_id = factor(model_id, levels=c("daily","daily manual","no da"))) |> 
    mutate(variable = ifelse(variable == "PHY_diatom", "diatom biomass", variable),
          variable = ifelse(variable == "PHY_cyano", "cyano biomass", variable),
         variable = ifelse(variable == "PHY_green", "green biomass", variable),
         variable = factor(variable, levels=c("temperature","oxygen","fdom", "chla", "green biomass", "cyano biomass", "diatom biomass"))) |> 
  filter(model_id %in% c("daily")) |> 
  ggplot(aes(x = datetime)) +
  geom_line(aes(y = prediction, group = parameter), color = "black") +
  geom_vline(aes(xintercept = as_datetime(reference_datetime)), color = "blue") +
  facet_wrap(~variable, scales = "free_y") +
  theme_bw() +
  labs(x = "date time", y = "forecast") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5,vjust = 0.5))
```

### Growth limitation

```{r}
open_dataset(relative_scores_directory) %>% 
  filter(horizon == 0, 
         (depth == 1.5 | is.na(depth)), 
         variable %in% c("PHY_green_fNit", "PHY_green_fPho", "PHY_green_fT",
                         "PHY_cyano_fNit", "PHY_cyano_fPho", "PHY_cyano_fT",
                         "PHY_diatom_fNit", "PHY_diatom_fPho", "PHY_diatom_fT")) %>% 
  collect() %>% 
    mutate(model_id = ifelse(model_id == "glmaed_ms_daily", "daily", model_id),
          model_id = ifelse(model_id == "glmaed_ms_daily_manual", "daily_manual", model_id),
         model_id = ifelse(model_id == "glmaed_ms_no_da", "no da", model_id),
         model_id = factor(model_id, levels=c("daily","daily_manual","no da"))) |> 
  select(datetime, variable, median, model_id) %>% 
  pivot_wider(names_from = variable, values_from = median) %>% 
  rowwise() %>% 
  mutate(PHY_green_growth = min(c(PHY_green_fNit, PHY_green_fPho)) *  PHY_green_fT * 2.0,
         PHY_cyano_growth = min(c(PHY_cyano_fNit, PHY_cyano_fPho)) *  PHY_cyano_fT * 1.0,
         PHY_diatom_growth = min(c(PHY_diatom_fNit, PHY_diatom_fPho)) *  PHY_diatom_fT * 3.0) %>% 
  pivot_longer(-c(datetime,model_id), names_to = "variable", values_to = "median") %>% 
  separate_wider_delim(cols = variable, delim = "_", names = c("PHY","phy_group","phy_variable")) |> 
  ggplot(aes(x = datetime, y = median, color = model_id)) +
  geom_line() +
  facet_grid(phy_variable~phy_group)
```


